Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-10-24 paper/main.tex

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common
representational space. Briefly, our approach uses a topic model~\citep[Latent
Dirichlet Allocation; ][]{BleiEtal03} trained on a set of documents, to
discover a set of (up to) $k$ ``topics'' or ``themes.'' Formally, each topic is
defined as a distribution of weights over words in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are semantically related, as implied by their co-occurring in the same
documents. After fitting a topic model, each document in the training set, or
any \textit{new} document that contains at least some of the words in the
model's vocabulary, may be represented as a $k$-dimensional vector describing
how much the document (most probably) reflects each topic. To select an
appropriate $k$ for our model, we identified the minimum number of topics that
yielded at least one ``unused'' topic (i.e., in which all words in the
vocabulary were assigned uniform weights) after training. This indicated that
the number of topics was sufficient to capture the set of latent themes present
in the two lectures (from which we constructed our document corpus, as
described below). We found this value to be $k = 15$ topics. The distribution
of weights over words in the vocabulary for each discovered topic is shown in
Supplementary Figure~\topicWordWeights, and each topic's top-weighted words may
be found in Supplementary Table~\topics.

>>>>

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common
representational space. Briefly, our approach uses a topic model~\citep[Latent
Dirichlet Allocation; ][]{BleiEtal03} trained on a set of documents, to
discover a set of (up to) $k$ ``topics'' or ``themes.'' Formally, each topic is
defined as a distribution of weights over words in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are semantically related, as implied by their co-occurring in the same
documents. After fitting a topic model, each document in the training set, or
any \textit{new} document that contains at least some of the words in the
model's vocabulary, may be represented as a $k$-dimensional vector describing
how much the document (most probably) reflects each topic. To select an
appropriate $k$ for our model, as a starting point, we identified the minimum
number of topics that yielded at least one ``unused'' topic (i.e., in which all
words in the vocabulary were assigned uniform weights) after training. This
indicated that the number of topics was sufficient to capture the set of latent
themes present in the two lectures (from which we constructed our document
corpus, as described below). We found this value to be $k = 15$ topics. We
found that with a limited number of additional adjustments
following~\citep{BoydMimn14}, such as removing corpus-specific stop-words, the
model yielded (subjectively) sensible and coherent topics. The distribution of
weights over words in the vocabulary for each discovered topic is shown in
Supplementary Figure~\topicWordWeights, and each topic's top-weighted words may
be found in Supplementary Table~\topics.
